---
title       : Bootstrapping
subtitle    : to estimate population statistics
author      : Mark Mo
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, mathjax, shiny, interactive]            # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: "libraries/polycharts"}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## What is bootstrapping?

_"To pull oneself up by one's bootstraps"_

* The seemingly impossible task of extracting more information from the same data set
* A goal of inferential statistics is to determine the value of a parameter of a population, such as the mean or median
* What if we can't make assumptions about the shape of the population distribution and only one sample exists
* Bootstrapping treats the original sample as a mini population from which to draw many (simulated) samples and therefore enable a confidence interval to be calculated
* However, the technique can be counterintuitive
* The aim of the [Bootstrap Demo App](https://markmo.shinyapps.io/bootstrapdemo/) is to help acquire the intuition

---

## Computer Science + Statistics = Data Science?

* Bootstrapping involves a relatively simple procedure, but repeated so many times that using a computer to perform the calculations is necessary
* A new sample of a given size is selected from the original sample with replacement
* This is repeated many times to generate a distribution of the sample statistic
* A confidence interval of the estimated parameter can be obtained from this distribution using the standard formula:

$$\bar x_b \pm z * SE_b$$

where \\(\bar x\\) is the mean of the bootstrap distribution, \\(z\\) is the relevant z-statistic for the confidence level, and \\(SE_b\\) is the standard error of the bootstrap distribution

---

## Bootstrapping example

* The following R code produces a 95% confidence interval for the estimate of the population mean from sample data using the bootstrapping technique:

```{r}
  mock_housing_prices <- sample(50000:100000, 300)
  sample_size <- 50; num_simulations <- 50
  sample_means <<- rep(NA, num_simulations)
  for (i in 1:num_simulations) {
    s <- sample(mock_housing_prices, sample_size, replace=TRUE)
    sample_means[i] <- mean(s)
  }
  boot_mean <- mean(sample_means)
  boot_se <- sd(sample_means) / sqrt(sample_size)
  boot_ci_95 <- boot_mean + c(-1, 1) * 1.96 * boot_se
  boot_ci_95
```

---

## A Histogram of the previous example

```{r r simple-plot, fig.height = 6, fig.align = 'center', message = F}
hist(sample_means)
```
